---
title: "Positive Leadership Behaviors Predict PERMA+4 in an African American Sample of Business Professionals"
author: "Scott Donaldson, PhD"
date: "10/04/21"
output: html_notebook
---

# Data Management

## Note from SPSS Upload

(N = 9616)

1.  If participants completed all scale variables they were retained

2.  Deleted 376

3.  Then deleted 10 for no to the informed consent question

## Upload Packages

```{r Load R Packages}
setwd("~/Desktop/USD/PositiveLeadership/R/PositiveLeadership")
library(haven)
library(tidyverse)
library(careless)
library(psych)
library(Hmisc)

```

## Read Data from SPSS

```{r Read dataset from SPSS}
# Read in data from Excel
PL_Data <- read_sav("PositiveLeadership.sav")

# Get rid of SPSS attributes and labels
PL_Data <- zap_label(PL_Data)
PL_Data <- zap_formats(PL_Data)
PL_Data <- zap_labels(PL_Data)
PL_Data <- zap_widths(PL_Data)
PL_Data <- as.data.frame(PL_Data)
PL_Data <- select(PL_Data, -Informed_Consent)
```

## Create Scale Scores

```{r Score Gill and PFW}
#Leadership Scale
PL_Data <- mutate(PL_Data,
LVM_Scale         = (LeadVisionMission1 + LeadVisionMission2)/2,
LStrat_Scale      = (LeadStrategy1 + LeadStrategy2)/2,
LEmpower_Scale    = (LeadEmpowerment1 + LeadEmpowerment2)/2,
LInfluence_Scale  = (LeadInfluence1 + LeadInfluence2)/2,
LShared_Scale     = (LeadShared1 + LeadShared2)/2,
GILL_Scale        = (LVM_Scale  + LStrat_Scale + LEmpower_Scale +
                     LInfluence_Scale + LShared_Scale)/5,

#Positive Functioning at Work
PL_Data <- mutate(PL_Data,
PE_Scale      = (Positive_Emotion_1 + Positive_Emotion_2)/2,
EN_Scale      = (Engagement_1 + Engagement_2)/2,
REL_Scale     = (Relationships_1 + Relationships_2)/2,
Mean_Scale    = (Meaning_1 + Meaning_2)/2,
Acc_Scale     = (Accomplishment_1 + Accomplishment_2)/2,
Health_Scale  = (Physical_Health_1 + Physical_Health_2)/2,
Mind_Scale    = (Mindset_1 + Mindset_2)/2,
ECON_Scale    = (Economic_Security_1 + Economic_Security_1)/2,
Enviro_Scale  = (Environment_1 + Environment_1)/2,

#Score Overall Scale
PFW_Scale     = (PE_Scale + 
                 EN_Scale +
                 REL_Scale +
                 Mean_Scale +
                 Acc_Scale +
                 Health_Scale +
                 Mind_Scale +
                 ECON_Scale +
                 Enviro_Scale)/9
))
      
```

## Recode demographics

```{r Demos 2 }
## Age
PL_Data$Age <- as.numeric(PL_Data$Age)

## Job Tenure
PL_Data$Job_Tenure <- as.numeric(PL_Data$Job_Tenure)

## Gender
PL_Data$Gender <- as.factor(PL_Data$Gender)
PL_Data$Gender <-fct_recode(PL_Data$Gender,
                    "Male"      = "9",
                    "Female"    = "10" ,
                    "Third"     = "11",
                    "PreferNA"  = "12",
                    "Other"     = "14"
                    )

## Ethnicity
PL_Data$Ethnicity <- as.factor(PL_Data$Ethnicity)
PL_Data$Ethnicity <-fct_recode(PL_Data$Ethnicity,
                    "White"    = "198",
                    "Black"    = "199" ,
                    "AIAN"     = "200",
                    "Asian"    = "201",
                    "NHPI"     = "202",
                    "Hispanic" = "203",
                    "mixed"    = "204",
                    "Other"    = "205",
                    )

## Industry
PL_Data$Current_Industry <- as.factor(PL_Data$Current_Industry)
PL_Data$Current_Industry <-fct_recode(PL_Data$Current_Industry,
                    "Banking"       = "4",
                    "Education"     = "5" ,
                    "Food&Bev"      = "6",
                    "Government"    = "7",
                    "Healthcare"    = "8",
                    "Manufacturing" = "9",
                    "Media&ENT"     = "10",
                    "Retail"        = "11",
                    "SoftwareIT"    = "12",
                    "Nonprofit"     = "13",
                    "Other"         = "14",
                    )

## Job Function
PL_Data$Job_Function <- as.factor(PL_Data$Job_Function)
PL_Data$Job_Function <-fct_recode(PL_Data$Job_Function,
                    "Accouting"      = "4",
                    "Administrative" = "5" ,
                    "Arts&Design"    = "6",
                    "Education"      = "7",
                    "Engineering"    = "8",
                    "IT"             = "9",
                    "Marketing"      = "10",
                    "Management"     = "11",
                    "Operations"     = "12",
                    "Other"          = "13"
                    )

## Income
PL_Data$Income <- as.factor(PL_Data$Income)
PL_Data$Income <-fct_recode(PL_Data$Income,
                    "Lessthan25k" = "9",
                    "25_50k"      = "10" ,
                    "50_75k"      = "11",
                    "75_99k"      = "13",
                    "100_150k"    = "14",
                    "150more"     = "20")

PL_Data$row_num <- seq.int(nrow(PL_Data)) 
```

# Data Cleaning

The first thing I will do is do a Mahalonobis Distance Outlier test on our two scales. Then I will examine longstring invariant responding to remove careless responders.

-   Findings:

    -   Removed 89 outliers based on Mahal

    -   Longstring showed an average of 5 and SD - 3.9. Thus, I decided to delete respondents that had 10 or more invariant responses.

        -   Remove 874 respondents

    -   Based on Desimone, less than 80 seconds should be deleted

        -   Delete 246

```{r Mahal}
MD <- mahad(PL_Data[, c(51,61)], 
            flag = TRUE, 
            confidence = 0.999) 

MD$row_num <- seq.int(nrow(MD)) 

PL_Data <- inner_join(PL_Data,
                      MD,
                      by = "row_num")
#Remove 89 outliers
PL_Data <- filter(PL_Data, flagged ==FALSE)
```

```{r Longstring}
Longstring <- as.data.frame(longstring(PL_Data[, c(3:33)]))
Longstring$row_num <- seq.int(nrow(Longstring)) 

#Boxplot plus descriptive stats
describe(Longstring)
boxplot(Longstring)

#Join the datasets
PL_Data$row_num <- seq.int(nrow(PL_Data))
PL_Data <- inner_join(PL_Data,
                      Longstring,
                      by = "row_num")
PL_Data <- filter(PL_Data, 
                  `longstring(PL_Data[, c(3:33)])` <=10)

```

```{r Duration in seconds}
describe(PL_Data$Duration__in_seconds_)
boxplot(PL_Data$Duration__in_seconds_)

PL_Data <- filter(PL_Data, 
                  Duration__in_seconds_ >= 80)

```

# Data Analysis
## Subset by African American
```{r}
AfriAmeri <- PL_Data[which(PL_Data$Ethnicity=='Black'), ]
```

## Demographics
```{r Demos}
# Age
mean(AfriAmeri$Age, na.rm = TRUE)
sd(AfriAmeri$Age, na.rm = TRUE)

# Gender
freqdist::freqdist(AfriAmeri$Gender)

# Industry
freqdist::freqdist(AfriAmeri$Current_Industry)

# Job Function
freqdist::freqdist(AfriAmeri$Job_Function)

# Income
freqdist::freqdist(AfriAmeri$Income)

# Job Tenure
mean(AfriAmeri$Job_Tenure, na.rm = TRUE)
sd(AfriAmeri$Job_Tenure, na.rm = TRUE)


```

## Descriptives, Cronbach, and Correlation Matrix
```{r Descriptives}
# Gill and PFW Means and SDs
sapply(AfriAmeri[, c(46:61)],
               mean)
sapply(AfriAmeri[, c(46:61)],
               sd)
# Alphas
alpha(AfriAmeri[, c(46:50)]) #Gill Scale
alpha(AfriAmeri[, c(52:60)]) #PFW

```
```{r Correlation Matrix}
Matrix <- AfriAmeri[, c(46:61)]
Matrix <- as.matrix(Matrix)
rcorr(Matrix, type = "pearson")

```

## Regression Modeling
```{r Create Dummy Codes for Gender & Ethnicity}
#Gender
Constrast.Gender <-contrasts(AfriAmeri$Gender) <- contr.treatment(5, 
                                               base = 1,
                                               contrasts = TRUE)




```
```{r GILL predicting PFW and subdimensions}
attach(AfriAmeri)

# GILL
Reg.GILL <- lm(PFW_Scale ~ GILL_Scale  + Gender)    
summary(Reg.GILL)
confint.lm(Reg.GILL)

# Vision and Mission
Reg.VM <- lm(PFW_Scale ~ LVM_Scale  + Gender)    
summary(Reg.VM)
confint.lm(Reg.VM)

# Strategy
Reg.Strat <- lm(PFW_Scale ~ LStrat_Scale  + Gender)    
summary(Reg.Strat)
confint.lm(Reg.Strat)

# Empowerment
Reg.Emp <- lm(PFW_Scale ~ LEmpower_Scale  + Gender)    
summary(Reg.Emp)
confint.lm(Reg.Emp)

# Influence
Reg.IMI <- lm(PFW_Scale ~ LInfluence_Scale  + Gender)    
summary(Reg.IMI)
confint.lm(Reg.IMI)

# Shared
Reg.Shared <- lm(PFW_Scale ~ LShared_Scale  + Gender)    
summary(Reg.Shared)
confint.lm(Reg.Shared)


```





## Exploratory Analyses
```{r Explore}
#Lump Gender
AfriAmeri$Gender <- fct_lump_n(AfriAmeri$Gender, 2)

# what does this data set say about fear?

##Boxplot of PFW by Gender
ggplot(AfriAmeri, aes(x=Gender, y=PFW_Scale)) + geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", shape=5, size=4)

##Boxplot of PFW by Income
ggplot(AfriAmeri, aes(x=Income, y=PFW_Scale)) + geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", shape=5, size=4)

## Income ANOVA
aov1 <- aov(PFW_Scale ~ Income, data=AfriAmeri)
summary(aov1)
model.tables(aov1, "means")

## Does gender x income make a difference
aov2 <- aov(PFW_Scale ~ Income*Gender, data=AfriAmeri)
aov2 <- aov(PFW_Scale ~ Income + Gender + Income:Gender, data=AfriAmeri)
summary(aov2)
model.tables(aov2, "means")

# Bar graph, time on x-axis, color fill grouped by sex -- use position_dodge()
ggplot(data=AfriAmeri, aes(x=Income, y=PFW_Scale, fill=Gender)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black")

##Boxplot of PFW by Industry
ggplot(AfriAmeri, aes(x=Job_Function, y=PFW_Scale)) + geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", shape=5, size=4)

# Facet wraps
sp <- ggplot(AfriAmeri, aes(x=GILL_Scale, y=PFW_Scale)) + geom_point(shape=1)
sp

sp + facet_grid(Income ~ .)
```

